---
title: "Interactive Event Prediction Report"
author: "Event ENM workflow"
date: "`r format(Sys.Date(), '%Y-%m-%d')`"
output:
  html_document:
    toc: true
    toc_float: true
    self_contained: false
runtime: shiny
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)

required_packages <- c("shiny", "leaflet", "terra", "htmltools")
missing_packages <- required_packages[!vapply(required_packages, requireNamespace, logical(1), quietly = TRUE)]
if (length(missing_packages) > 0) {
  stop("Install required packages before running this report: ", paste(missing_packages, collapse = ", "), call. = FALSE)
}

library(shiny)
library(leaflet)
library(htmltools)

get_current_script_path <- function() {
  command_args <- commandArgs(trailingOnly = FALSE)
  file_arg <- grep("^--file=", command_args, value = TRUE)
  if (length(file_arg) > 0) {
    return(normalizePath(sub("^--file=", "", file_arg[1]), winslash = "/", mustWork = TRUE))
  }

  for (frame in rev(sys.frames())) {
    if (!is.null(frame$ofile)) {
      return(normalizePath(frame$ofile, winslash = "/", mustWork = TRUE))
    }
  }

  if (requireNamespace("rstudioapi", quietly = TRUE) && rstudioapi::isAvailable()) {
    active_doc <- tryCatch(rstudioapi::getActiveDocumentContext()$path, error = function(e) NA_character_)
    if (!is.na(active_doc) && nzchar(active_doc)) {
      return(normalizePath(active_doc, winslash = "/", mustWork = TRUE))
    }
  }

  NA_character_
}

find_code_dir <- function() {
  script_path <- get_current_script_path()
  wd <- normalizePath(getwd(), winslash = "/", mustWork = FALSE)
  script_dir <- if (!is.na(script_path)) dirname(script_path) else NA_character_
  wd_parents <- normalizePath(
    file.path(wd, c(".", "..", "../..", "../../..")),
    winslash = "/",
    mustWork = FALSE
  )
  script_candidates <- if (!is.na(script_dir)) {
    normalizePath(
      file.path(script_dir, c(".", "..", "../..")),
      winslash = "/",
      mustWork = FALSE
    )
  } else {
    character(0)
  }

  candidates <- c(
    script_candidates,
    wd_parents,
    file.path(wd_parents, "KSPH Code")
  )

  candidates <- unique(normalizePath(candidates, winslash = "/", mustWork = FALSE))
  for (candidate in candidates) {
    if (
      file.exists(file.path(candidate, "config", "predictor_list.csv")) &&
      dir.exists(file.path(candidate, "R_python_code"))
    ) {
      return(candidate)
    }
  }

  stop(
    "Could not locate the KSPH Code repo root. Open this report from inside the KSPH Code project, ",
    "or set CODE_DIR manually near the top of 04_interactive_prediction_report_Shiny.Rmd.",
    call. = FALSE
  )
}

CODE_DIR <- find_code_dir()
PREDICTION_DIR <- file.path(CODE_DIR, "outputs", "predictions")
TEMPORAL_FORWARD_DIR <- file.path(CODE_DIR, "outputs", "tfv", "all_types")
TEMPORAL_FORWARD_COMPARISON_PREDICTION_DIR <- file.path(
  TEMPORAL_FORWARD_DIR,
  "maps_2022_2025",
  "pred"
)

PREDICTION_YEARS <- 2020:2025
CHANGE_YEARS <- 2021:2025
TEMPORAL_FORWARD_PREDICTION_YEARS <- 2022:2025
TEMPORAL_FORWARD_CHANGE_YEARS <- 2023:2025
RASTER_CRS <- "EPSG:4326"
TOP_PERCENTILE <- 0.99
MAP_OPACITY_VALUE <- 0.72
MAP_OPACITY_TOP1 <- 0.92
LEAFLET_MAP_HEIGHT <- "300px"
LEAFLET_NA_COLOR <- "#00000000"

ROR_BREAKPOINTS <- c(0.4, 0.7, 1.5, 5, 15)
CHANGE_RATIO_BREAKPOINTS <- c(0.6, 0.8, 1.2, 1.6, 2)
ROR_VALUE_COLORS <- c("green4", "lightgreen", "white", "red", "darkred", "purple4")
CHANGE_RATIO_VALUE_COLORS <- c("blue3", "lightblue", "white", "orange1", "red2", "red4")
TOP_PERCENTILE_COLOR <- "red"

raster_cache <- new.env(parent = emptyenv())

temporal_forward_run_name <- function(year) {
  sprintf("tr2001_%s_p%s", as.integer(year) - 1L, as.integer(year))
}

temporal_forward_ror_path <- function(year) {
  comparison_path <- file.path(
    TEMPORAL_FORWARD_COMPARISON_PREDICTION_DIR,
    "ror",
    sprintf("event_probability_ROR_%s.tif", year)
  )
  run_path <- file.path(
    TEMPORAL_FORWARD_DIR,
    temporal_forward_run_name(year),
    "pred",
    "ror",
    sprintf("event_probability_ROR_%s.tif", year)
  )

  if (file.exists(comparison_path) || !file.exists(run_path)) {
    return(comparison_path)
  }
  run_path
}

prediction_sources <- list(
  all_years = list(
    label = "Fully trained model",
    prediction_years = PREDICTION_YEARS,
    change_years = CHANGE_YEARS,
    ror_path = function(year) {
      file.path(PREDICTION_DIR, "ROR estimates", sprintf("event_probability_ROR_%s.tif", year))
    },
    ror_change_path = function(year) {
      file.path(PREDICTION_DIR, "ROR 1yr ratios", sprintf("chgROR_1yr_%s_over_%s.tif", year, year - 1))
    }
  ),
  step_forward = list(
    label = "Step-forward model",
    prediction_years = TEMPORAL_FORWARD_PREDICTION_YEARS,
    change_years = TEMPORAL_FORWARD_CHANGE_YEARS,
    ror_path = temporal_forward_ror_path,
    ror_change_path = function(year) {
      file.path(
        TEMPORAL_FORWARD_COMPARISON_PREDICTION_DIR,
        "ror_1yr",
        sprintf("chgROR_1yr_%s_over_%s.tif", year, year - 1)
      )
    }
  )
)

event_report_products <- list(
  ror = list(
    label = "Relative odds ratio",
    value_choice_label = "Overall ROR predictions",
    top_choice_label = "Top 1% of ROR predictions",
    source_years = function(source) source$prediction_years,
    year_label = function(year) as.character(year),
    raster_path = function(source, year) source$ror_path(year),
    value_layer = "ROR_mean",
    top_layer = "ROR_mean_top1pct",
    breakpoints = ROR_BREAKPOINTS,
    colors = ROR_VALUE_COLORS,
    legend_title = "Mean ROR",
    top_legend_title = "Top 1% ROR"
  ),
  ror_change = list(
    label = "1-year ROR ratio",
    value_choice_label = "1-year ratio measure",
    top_choice_label = "Top 1% of 1-year ratios",
    source_years = function(source) source$change_years,
    year_label = function(year) sprintf("%s / %s", year, year - 1),
    raster_path = function(source, year) source$ror_change_path(year),
    value_layer = "chgROR_1yr_mean",
    top_layer = "chgROR_1yr_mean_top1pct",
    breakpoints = CHANGE_RATIO_BREAKPOINTS,
    colors = CHANGE_RATIO_VALUE_COLORS,
    legend_title = "Mean ratio",
    top_legend_title = "Top 1% ratios"
  )
)

available_year_choices <- function(product, source_id) {
  source <- prediction_sources[[source_id]]
  if (is.null(source)) {
    return(character(0))
  }

  product_years <- product$source_years(source)
  existing <- product_years[vapply(
    product_years,
    function(year) file.exists(product$raster_path(source, year)),
    logical(1)
  )]
  if (length(existing) == 0) {
    return(character(0))
  }

  stats::setNames(as.character(existing), vapply(existing, product$year_label, character(1)))
}

available_source_choices <- function(product_id) {
  product <- event_report_products[[product_id]]
  if (is.null(product)) {
    return(character(0))
  }

  source_ids <- names(prediction_sources)
  source_ids <- source_ids[vapply(
    source_ids,
    function(source_id) length(available_year_choices(product, source_id)) > 0,
    logical(1)
  )]

  stats::setNames(source_ids, vapply(source_ids, function(source_id) prediction_sources[[source_id]]$label, character(1)))
}

default_source_choice <- function(product_id) {
  choices <- available_source_choices(product_id)
  if (length(choices) == 0) {
    return(NA_character_)
  }
  if ("all_years" %in% unname(choices)) {
    return("all_years")
  }
  unname(choices)[1]
}

available_product_sources <- lapply(names(event_report_products), available_source_choices)
names(available_product_sources) <- names(event_report_products)
missing_products <- names(available_product_sources)[vapply(available_product_sources, length, integer(1)) == 0]
if (length(missing_products) > 0) {
  stop(
    "No raster files were found for report product(s): ",
    paste(missing_products, collapse = ", "),
    ". Run the derived raster sections in 03_train_predict_brt_simple.R and/or Section 5 of 03b_temporal_forward_validation_brt.R first.",
    call. = FALSE
  )
}

pretty_legend_value <- function(value) {
  formatted <- format(signif(value, 3), scientific = FALSE, trim = TRUE)
  sub("\\.?0+$", "", formatted)
}

discrete_break_labels <- function(breaks) {
  if (length(breaks) < 2) {
    return(character(0))
  }

  labels <- character(length(breaks) - 1)
  for (i in seq_along(labels)) {
    lower <- pretty_legend_value(breaks[i])
    upper <- pretty_legend_value(breaks[i + 1])
    if (i == 1) {
      labels[i] <- paste0("<", upper)
    } else if (i == length(labels)) {
      labels[i] <- paste0(lower, "+")
    } else {
      labels[i] <- paste0(lower, "-", upper)
    }
  }
  labels
}

value_bins <- function(raster_layer, breakpoints) {
  values <- raster_numeric_values(raster_layer)
  values <- values[is.finite(values)]
  if (length(values) == 0) {
    return(c(0, 1))
  }

  min_value <- min(values, na.rm = TRUE)
  max_value <- max(values, na.rm = TRUE)
  if (min_value == max_value) {
    pad <- max(abs(min_value) * 0.05, 1e-6)
    return(c(min_value - pad, max_value + pad))
  }

  breaks <- sort(unique(c(
    min_value,
    breakpoints[breakpoints > min_value & breakpoints < max_value],
    max_value
  )))
  if (length(breaks) < 2) {
    pad <- max(abs(min_value) * 0.05, 1e-6)
    breaks <- c(min_value - pad, max_value + pad)
  }

  breaks
}

raster_numeric_values <- function(raster_layer) {
  raw_values <- terra::values(raster_layer)
  if (is.data.frame(raw_values)) {
    raw_values <- raw_values[[1]]
  } else if (is.matrix(raw_values)) {
    raw_values <- raw_values[, 1]
  } else if (is.list(raw_values)) {
    raw_values <- raw_values[[1]]
  }

  suppressWarnings(as.numeric(raw_values))
}

layer_values <- function(raster_layer) {
  values <- raster_numeric_values(raster_layer)
  values[is.finite(values)]
}

cache_key <- function(source_id, product_id, year) {
  paste(source_id, product_id, year, sep = "_")
}

read_product_raster <- function(source_id, product_id, year) {
  key <- cache_key(source_id, product_id, year)
  if (exists(key, envir = raster_cache, inherits = FALSE)) {
    return(get(key, envir = raster_cache, inherits = FALSE))
  }

  source <- prediction_sources[[source_id]]
  if (is.null(source)) {
    stop("Unknown prediction source: ", source_id, call. = FALSE)
  }
  product <- event_report_products[[product_id]]
  raster_path <- product$raster_path(source, year)
  if (!file.exists(raster_path)) {
    stop("Raster file does not exist: ", raster_path, call. = FALSE)
  }

  raster_stack <- terra::rast(raster_path)
  raster_crs <- terra::crs(raster_stack)
  if (is.na(raster_crs) || !nzchar(raster_crs)) {
    terra::crs(raster_stack) <- RASTER_CRS
  }

  assign(key, raster_stack, envir = raster_cache)
  raster_stack
}

make_top_percentile_layer <- function(value_layer, top_layer_name) {
  values <- layer_values(value_layer)
  if (length(values) == 0) {
    top_layer <- value_layer * NA_real_
  } else {
    threshold <- as.numeric(stats::quantile(values, probs = TOP_PERCENTILE, na.rm = TRUE, names = FALSE))
    top_layer <- terra::ifel(value_layer >= threshold, 1, NA_real_)
  }

  names(top_layer) <- top_layer_name
  top_layer
}

read_report_layer <- function(source_id, product_id, year, display_mode) {
  source <- prediction_sources[[source_id]]
  if (is.null(source)) {
    stop("Unknown prediction source: ", source_id, call. = FALSE)
  }
  product <- event_report_products[[product_id]]
  raster_stack <- read_product_raster(source_id, product_id, year)

  if (display_mode == "value") {
    if (!product$value_layer %in% names(raster_stack)) {
      stop("Layer '", product$value_layer, "' was not found in ", product$raster_path(source, year), call. = FALSE)
    }
    return(raster_stack[[product$value_layer]])
  }

  if (product$top_layer %in% names(raster_stack)) {
    top_layer <- raster_stack[[product$top_layer]]
    top_layer <- terra::ifel(top_layer >= 1, 1, NA_real_)
    names(top_layer) <- product$top_layer
    return(top_layer)
  }

  if (!product$value_layer %in% names(raster_stack)) {
    stop("Layer '", product$value_layer, "' was not found in ", product$raster_path(source, year), call. = FALSE)
  }
  make_top_percentile_layer(raster_stack[[product$value_layer]], product$top_layer)
}

raster_layer_to_cells <- function(raster_layer) {
  raster_layer <- terra::rast(raster_layer)
  values <- raster_numeric_values(raster_layer)
  finite_cells <- which(is.finite(values))
  if (length(finite_cells) == 0) {
    return(data.frame())
  }

  xy <- terra::xyFromCell(raster_layer, finite_cells)

  cell_resolution <- terra::res(raster_layer)
  data.frame(
    lng1 = xy[, "x"] - cell_resolution[1] / 2,
    lat1 = xy[, "y"] - cell_resolution[2] / 2,
    lng2 = xy[, "x"] + cell_resolution[1] / 2,
    lat2 = xy[, "y"] + cell_resolution[2] / 2,
    value = values[finite_cells]
  )
}

raster_value_summary <- function(raster_layer) {
  values <- raster_numeric_values(raster_layer)
  finite_values <- values[is.finite(values)]
  if (length(finite_values) == 0) {
    return("finite cells: 0")
  }

  paste0(
    "finite cells: ", format(length(finite_values), big.mark = ","),
    "; range: ", pretty_legend_value(min(finite_values, na.rm = TRUE)),
    " to ", pretty_legend_value(max(finite_values, na.rm = TRUE))
  )
}

fit_raster_bounds <- function(map, raster_layer) {
  extent <- terra::ext(raster_layer)
  leaflet::fitBounds(
    map,
    lng1 = terra::xmin(extent),
    lat1 = terra::ymin(extent),
    lng2 = terra::xmax(extent),
    lat2 = terra::ymax(extent)
  )
}

make_report_map <- function(source_id, product_id, year, display_mode) {
  source <- prediction_sources[[source_id]]
  if (is.null(source)) {
    stop("Unknown prediction source: ", source_id, call. = FALSE)
  }
  product <- event_report_products[[product_id]]
  raster_layer <- read_report_layer(source_id, product_id, year, display_mode)
  finite_values <- layer_values(raster_layer)

  map <- leaflet::leaflet(options = leaflet::leafletOptions(preferCanvas = TRUE)) |>
    leaflet::addProviderTiles(leaflet::providers$Esri.WorldTopoMap, group = "Topo") |>
    leaflet::addProviderTiles(leaflet::providers$CartoDB.Positron, group = "Light") |>
    leaflet::addScaleBar(position = "bottomleft")

  if (length(finite_values) == 0) {
    return(
      map |>
        leaflet::addControl(
          html = htmltools::HTML(paste0(
            "<strong>No finite raster values were found.</strong><br>",
            "Product: ", htmltools::htmlEscape(product$label), "<br>",
            "Source: ", htmltools::htmlEscape(source$label), "<br>",
            "Year: ", htmltools::htmlEscape(product$year_label(year)), "<br>",
            "Layer: ", htmltools::htmlEscape(names(raster_layer)[1]), "<br>",
            htmltools::htmlEscape(raster_value_summary(raster_layer))
          )),
          position = "topright"
        ) |>
        fit_raster_bounds(raster_layer)
    )
  }

  if (display_mode == "value") {
    bins <- value_bins(raster_layer, product$breakpoints)
    legend_labels <- discrete_break_labels(bins)
    colors <- grDevices::colorRampPalette(product$colors)(length(legend_labels))
    palette <- leaflet::colorBin(
      palette = colors,
      bins = bins,
      domain = finite_values,
      na.color = LEAFLET_NA_COLOR
    )

    map <- map |>
      leaflet::addRasterImage(
        raster_layer,
        colors = palette,
        opacity = MAP_OPACITY_VALUE,
        project = TRUE,
        maxBytes = 50 * 1024 * 1024,
        group = product$value_choice_label
      ) |>
      leaflet::addLegend(
        position = "bottomright",
        colors = rev(colors),
        labels = rev(legend_labels),
        title = product$legend_title,
        opacity = MAP_OPACITY_VALUE
      )
  } else {
    palette <- leaflet::colorFactor(
      palette = TOP_PERCENTILE_COLOR,
      domain = 1,
      na.color = LEAFLET_NA_COLOR
    )

    map <- map |>
      leaflet::addRasterImage(
        raster_layer,
        colors = palette,
        opacity = MAP_OPACITY_TOP1,
        project = TRUE,
        maxBytes = 50 * 1024 * 1024,
        group = product$top_choice_label
      ) |>
      leaflet::addLegend(
        position = "bottomright",
        colors = TOP_PERCENTILE_COLOR,
        labels = "Top 1%",
        title = product$top_legend_title,
        opacity = MAP_OPACITY_TOP1
      )
  }

  map |>
    leaflet::addLayersControl(
      baseGroups = c("Topo", "Light"),
      options = leaflet::layersControlOptions(collapsed = TRUE)
    ) |>
    fit_raster_bounds(raster_layer)
}

year_select_control <- function(product_id, source_id, year_input_id) {
  product <- event_report_products[[product_id]]
  choices <- available_year_choices(product, source_id)

  if (length(choices) == 0) {
    return(htmltools::div(class = "missing-source-note", "No rasters found for this source."))
  }

  shiny::selectInput(
    inputId = year_input_id,
    label = "Year",
    choices = choices,
    selected = choices[[length(choices)]]
  )
}

section_controls <- function(product_id, source_input_id, year_output_id, display_input_id) {
  product <- event_report_products[[product_id]]
  source_choices <- available_source_choices(product_id)
  if (length(source_choices) == 0) {
    return(htmltools::div(class = "missing-source-note", "No rasters found for this map section."))
  }

  htmltools::tagList(
    htmltools::div(
      class = "map-controls",
      shiny::fluidRow(
        shiny::column(
          width = 4,
          shiny::selectInput(
            inputId = source_input_id,
            label = "Prediction source",
            choices = source_choices,
            selected = default_source_choice(product_id)
          )
        ),
        shiny::column(
          width = 3,
          shiny::uiOutput(year_output_id)
        ),
        shiny::column(
          width = 5,
          shiny::selectInput(
            inputId = display_input_id,
            label = "Layer",
            choices = stats::setNames(
              c("value", "top"),
              c(product$value_choice_label, product$top_choice_label)
            ),
            selected = "value"
          )
        )
      )
    )
  )
}
```

```{css, echo=FALSE}
.main-container {
  max-width: 1600px !important;
  width: 96% !important;
}

.section-note {
  max-width: 1100px;
  font-size: 1.05rem;
  line-height: 1.45;
  margin-bottom: 1rem;
}

.map-controls {
  margin: 0.75rem 0 1rem 0;
}

.missing-source-note {
  color: #8a4b00;
  font-weight: 600;
  margin: 0.75rem 0;
}

.leaflet {
  border: 1px solid #cccccc;
}
```

This report reads the GeoTIFF rasters generated by the fully trained `03_train_predict_brt_simple.R` workflow and, when available, the step-forward `03b_temporal_forward_validation_brt.R` workflow. Use the prediction-source selector to switch between the full model and the temporally forward-stepped model outputs. Run it with `rmarkdown::run("R_python_code/04_interactive_prediction_report_Shiny.Rmd")` from the KSPH Code repo root so the selectors remain active.

## Overall Relative Odds Ratio Predictions

<div class="section-note">
These maps show the mean relative odds ratio (ROR) for each prediction-grid cell. A value above 1 means the estimated odds for that cell-year are higher than the study-area average odds for the same year. The top 1% option highlights cells with the highest mean ROR values for the selected year.
</div>

```{r ror-ui}
section_controls("ror", "ror_source", "ror_year_ui", "ror_display")
leaflet::leafletOutput("ror_map", width = "100%", height = LEAFLET_MAP_HEIGHT)
```

```{r ror-server}
output$ror_year_ui <- shiny::renderUI({
  req(input$ror_source)
  year_select_control("ror", input$ror_source, "ror_year")
})

output$ror_map <- leaflet::renderLeaflet({
  req(input$ror_source, input$ror_year, input$ror_display)
  make_report_map(input$ror_source, "ror", as.integer(input$ror_year), input$ror_display)
})
```

## 1-Year Change In Relative Odds Ratios

<div class="section-note">
These maps show the ratio of the current year's mean ROR to the previous year's mean ROR for the same cell. Values above 1 indicate an increase relative to the prior year; values below 1 indicate a decrease. The top 1% option highlights cells with the largest one-year increases for the selected year pair.
</div>

```{r ror-change-ui}
section_controls("ror_change", "ror_change_source", "ror_change_year_ui", "ror_change_display")
leaflet::leafletOutput("ror_change_map", width = "100%", height = LEAFLET_MAP_HEIGHT)
```

```{r ror-change-server}
output$ror_change_year_ui <- shiny::renderUI({
  req(input$ror_change_source)
  year_select_control("ror_change", input$ror_change_source, "ror_change_year")
})

output$ror_change_map <- leaflet::renderLeaflet({
  req(input$ror_change_source, input$ror_change_year, input$ror_change_display)
  make_report_map(input$ror_change_source, "ror_change", as.integer(input$ror_change_year), input$ror_change_display)
})
```
